---
title: "The DA provides some closing arguments"
author: "c'mon son"
date: "April 20, 2016"
output: pdf_document
---

1. Good on names. Do we need all of that output from before the problem?

2. All of the "eat" variables appear to be positively correlated, in varying magnitude. "bodyfat", "guilty", and "thinner" appear to be grouped together with very high correlations between them. All three variables could be viewed as negative towards eating and fat, hence the trait we call "negativity". Potentially another grouping of variables exist between "exercise", "avoid", and "terrified". In fact, we would consider the variable "empty" also shares some trait with these variables. This group of variables are moderately correlated with each other. They also have moderate to strong correlations with the previous group of variables, which suggests the variables may share some negativity towards eating and fat. These three variables may share a trait "coping". Each question associated with these variables may have triggered people to put into action some response to their feelings about food, either healthy or unhealthy. At the bottom of the plot, the variables "preoccupied" and "controls" are very highly correlated with each other and weak to moderately correlated with the other variables. The variable "time" also appears very correlated with these two. This set of variables are all have a food dominated relationship, hence they share the trait we call "domination". (__Maybe one of the names should be perception and one be anxiety.__)

3. To paraphrase from the help information, __mat.sort()__ attempts to sory things by having similar items grouped together. It does this through factor analysis, sorting by size of the largest factor loading for each variable. The default method sorts by the loadings on the first factor.

4. Looks like the variables were sorted more "correctly". "empty" and "time" were flipped which makes sense, as were "terrified" and "exercise". Similar correlations are sen though as evidenced by none of the differences in correlations being greater than 0.0882. The pearson correlations were typically smaller, but a few were larger.

```{r echo=T, eval=F}
require(polycor)
edo<-apply(renamed[,-c(1,12:14)],2,ordered)
polyc <- hetcor(edo)
pearc <- cor(renamed[,-c(1,12:14)])
max(as.matrix(polyc) - pearc)
cor.plot(mat.sort(as.matrix(polyc)))
```

5. Change the name of the table? I don't think we need to say anything...

6. Can we get the numbers to show up? Again, no words appear necessary...

7. In testing $H_0:$ the model is adequate vs. $H_A:$ something more complex is needed, a goodness-of-fit test results in $\chi^2=114.706$ test-statistic value that follows a $\chi^2_{34}$ distribution under the null hypothesis. This results in a p-value < 0.001 which is very strong evidence that a model that is more complex is needed here. We also see that the model has an RMSEA (Root Mean Square Error of Approximation) of 0.099 (an approximate 90% confidence interval being (0.079, 0.119)) which is larger than the desirable 0.05 and again suggests the model is not a good fit. 

8. All of the loadings (from the latent variable table) have very small p-values, suggesting strong evidence of their inclusion (all lines were testing $H_0: \lambda_i = 0$ vs. $H_A: \lambda_i \neq = 0$) despite the model not fitting well. __SHOULD WE BE TALKING ABOUT THE UNIQUENESSES?__

9. The disattenuated correlation of the latent traits, food and nonfood, is 0.65. This suggests that there is quite a bit of shared information being explained by both these latent traits. Due to the nature of all the questions having some affiliation with eating, this seems to make sense. __MORE PROBABLY SHOULD BE DISCUSSED, BUT NOT SURE WHAT TO SAY YET.__ I wasn't sure if he said something about correlations between latent traits.

10. Polychoric correlations were fit between two ordered categorical variables. Pearson correlations were fit between two quantitative variables. Polyserial correlations were fit between an ordered categorical variable and a quantitative variable. As an update to the previous discussion of the correlation plot, there are weak to moderate positive correlations between any of the quantitative variables and any other variables.

11. I think you fit endogenous variables and not exogenous. Arrows should be going out, not in. I think it should be (pg 25 of CFA and SEM notes): maybe not based on a website I read....

```{r p11_a, eval=F}
all.cfa <- cfa(model, sample.cov = polyc$correlations, sample.nobs = 242)
sem.model <- paste(model, "\n food ~ bmi
                        nonfood ~ bmi
                        food ~ wsb 
                        nonfood ~ wsb
                        food ~ anx
                        nonfood ~ anx")
sem.fit <- sem(sem.model, sample.cov = allc$correlations, sample.nobs = 242, std.lv=T)
pander(parTable(sem.fit))
summary(sem.fit, fit.measures=T)
semPaths(sem.fit, "par")
```

What summary do we report? Do we need to discuss anything? GOF is still suggesting poor fit, do we care? AIC went up about 2000, does that matter? I mean does it make sense to compare a CFA to an SEM? Do we just give output and not discuss anything?

12. Both tests have small p-values (<0.001) suggesting that bmi is helping to explain both variables.

```{r p12_a, eval=F}
summary(sem.fit, fit.measures=T)
```

If instead, you are correct in the order of the formula, the test for bmi involving food has a p-value of 0.076 which is moderate to weak evidence and the test for bmi involving nonfood has a p-value of 0.002 which is strong evidence. Are we suppose to say bmi causes nonfood or vice versa?

13. Trying to explain underlying latent traits using the observed variables.???? It is trying to establish causality of latent variables.