---
title: "STAT 537 HW 9"
author: "Doug Anderson and Brandon Fenton"
date: "Due April 21, 2016"
output:
  pdf_document: default
  fig_width: 8
  fig_height: 5
  word_document: null
---

You can works in groups of up to 3, no restrictions on composition.

You can find the survey used to generate the "eat" labeled questions you will be working with here: https://psychology-tools.com/eat-26/

The numbers on the "eat" questions are the same as in the online survey although the version you have used a 7-point Likert scale (with 7 a ) instead of the 6 point version on the website. bmi is a version of the body mass index, wsb is a score on western standards of beauty with (I assume) higher scores relating to higher self-perceived beauty of the subject, and anx is an anxiety score with (I assume) higher scores relating to higher anxiety.

Note that there are some missing values in the data set that were coded as -99. The following code will get you going on this but is not complete:

```{r setup,warning=F,message=F,fig.width=10, fig.height=8, echo=F}
# eatingatt_data<-read.csv("https://montana.box.com/shared/static/ihwz8fjd024a5e8zuxzwgs90q2zhxpmc.csv",na.strings=-99)

eatingatt_data <- read.csv("eating.csv")
ead_c<-na.omit(eatingatt_data) #Cleaning up the data set to allow focus on other aspects - could retain nearly 60 partial observations 
require(psych)
describe(ead_c)

#Rename eat variables and then proceed with the following...

cor.plot(mat.sort(cor(ead_c[,-c(1,12:14)])))

require(lavaan)
require(polycor)
require(semPlot)

edo<-apply(ead_c[,-c(1,12:14)],2,ordered)

require(pander)
knitr::opts_chunk$set(echo=F, warning=F, message=F, cache=T, comment=NA, size="footnotesize", fig.width=6, fig.height=4, fig.align="center", dev="pdf", dev.args=list(pointsize=8))
panderOptions("round", 5)


```

1) First make a table of the questions being used here and rename each "eat" variable to make it more explicit about what the question was focused on. 

```{r p1_a}
questions <- c("I am terrified about being overweight.",
"I avoid eating when I am hungry.",
"I feel extremely guilty after eating.",
"I am occupied with a desire to be thinner.",
"I think about burning up calories when I exercise.",
"I am preoccupied with the thought of having fat on my body.",
"I like my stomach to be empty.",
"I find myself preoccupied with food.",
"I feel that food controls my life.",
"I give too much time and thought to food.")

new.names <-c("terrified",
"avoid",
"guilty",
"thinner",
"exercise",
"bodyfat",
"empty",
"preoccupied",
"controls",
"time")

q.table <- data.frame(questions, colnames(ead_c)[2:11], new.names)
colnames(q.table) <- c("Questions", "Old name", "New Name")

pander(q.table)
renamed <- ead_c
colnames(renamed)[2:11] <- new.names
```


Using your new names re-run the provided cor.plot code (you do not need to report the previous plot). I had some plot margins too large issues in compiling in markdown with this plot until I changed the figure size. You can always do it in outside of markdown and load the figure back into the final compile if these options don't work for you. Then: 

2) Discuss the correlation matrix pattern.

```{r p2_a, fig.width=10, fig.height=8}
cor.plot(mat.sort(cor(renamed[,-c(1,12:14)])))
```

All of the "eat" variables appear to be positively correlated, in varying magnitude. "bodyfat", "guilty", and "thinner" appear to be grouped together with very high correlations between them. All three variables could be viewed as negative towards eating and fat, hence the trait we call "negativity". Potentially another grouping of variables exist between "exercise", "avoid", and "terrified". In fact, we would consider the variable "empty" also shares some trait with these variables. This group of variables are moderately correlated with each other. They also have moderate to strong correlations with the previous group of variables, which suggests the variables may share some negativity towards eating and fat. These three variables may share a trait "coping". Each question associated with these variables may have triggered people to put into action some response to their feelings about food, either healthy or unhealthy. At the bottom of the plot, the variables "preoccupied" and "controls" are very highly correlated with each other and weak to moderately correlated with the other variables. The variable "time" also appears very correlated with these two. This set of variables are all have a food dominated relationship, hence they share the trait we call "domination". (__Maybe one of the names should be perception and one be anxiety.__)

3) Use the help information to explain how mat.sort() sorted the correlation matrix.

To paraphrase from the help information, __mat.sort()__ attempts to sory things by having similar items grouped together. It does this through factor analysis, sorting by size of the largest factor loading for each variable. The default method sorts by the loadings on the first factor.

4) The previous result was using Pearson correlation. Identify a more appropriate correlation measure for these data (report it). Then estimate the correlation matrix using functions from the polycor package (not psych) and compare the correlations to those produced by Pearson correlation here, comparing the estimates and remaking the correlation matrix plot.

Looks like the variables were sorted more "correctly". "empty" and "time" were flipped which makes sense, as were "terrified" and "exercise". Similar correlations are sen though as evidenced by none of the differences in correlations being greater than 0.0882. The pearson correlations were typically smaller, but a few were larger.

```{r p4_a}
edo<-apply(renamed[,-c(1,12:14)],2,ordered)
polyc <- hetcor(edo)
pearc <- cor(renamed[,-c(1,12:14)])
max(as.matrix(polyc) - pearc)
cor.plot(mat.sort(as.matrix(polyc)))
```

5) Using your better named variables and your "better" correlation matrix, fit a CFA that puts eat21, eat3, and eat18 together on one latent trait and the other eat variables on the other. To make sure it is using your correlation matrix, use sample.cov and sample.nobs and do not provide the data set to the cfa function from lavaan. Report the summary output from the model with the latent trait variances fixed at one.

```{r p5_a}
# eat21: I give too much time and thought to food.
# eat3: I find myself preoccupied with food.
# eat18: I feel that food controls my life.

model <- 'food =~ time + controls + preoccupied
          nonfood =~ empty + bodyfat + exercise + thinner + guilty + avoid + terrified'
mcfa <- cfa(model, sample.cov = polyc$correlations, sample.nobs = 242, std.lv = T)
pander(parTable(mcfa))
```

6) Make a path diagram of the resulting model.

```{r p6_a}
require(semPlot)
semPaths(mcfa, "par")
```

7) Discuss the quality of the fit of the model using (a) the goodness of fit test and (b) one of the other measures of quality of fit. Make sure to fully report the numerical results as well as discussing what this tells you about the model.

```{r p7_a}
# TODO There's got to be a better way! Most likely involving fitMeasures() 
summary(mcfa, fit.measures=T)
```

8) Discuss all the loadings - are there any that potentially could be dropped? 

9) Report and discuss the disattenuated correlation of the latent traits. Does this make sense in this application?

10) Add the other three variables to your correlation matrix using appropriate correlation measures. Note the types of correlations you are using (you can discuss this generally). Also update your correlation matrix plot using this matrix.

```{r p10_a}
eda <- cbind(edo, renamed[,12:14])
allc <- hetcor(eda)
cor.plot(mat.sort(as.matrix(allc)))
```

11) Fit an SEM that retains the same latent variable construction as in the CFA but that also uses bmi, anx, and wsb as exogenous variables that explain both of the latent traits. Fit the model and report the summary. Also make a path diagram.

```{r p11_a}
all.cfa <- cfa(model, sample.cov = polyc$correlations, sample.nobs = 242)
sem.model <- paste(model, "\n bmi ~ food + nonfood
                        wsb ~ food + nonfood
                        anx ~ food + nonfood")
sem.fit <- sem(sem.model, sample.cov = allc$correlations, sample.nobs = 242, std.lv=T)
semPaths(sem.fit, "par")
```

12) Report the evidence and discuss the interpretation of the two tests for the paths that relate bmi to each of the latent traits. 

```{r p12_a}
summary(sem.fit, fit.measures=T)
```

13) Generally explain what this model is exploring.


## R Code Appendix:
Problem 1:
```{r a1, ref.label='p1_a', eval=F, echo=T}
```

```{r b1, ref.label='p1_b', eval=F, echo=T}
```

Problem 2:
```{r a2, ref.label='p2_a', eval=F, echo=T}
```

\pagebreak
Problem 4:
```{r a4, ref.label='p4_a', eval=F, echo=T}
```

Problem 5:
```{r a5, ref.label='p5_a', eval=F, echo=T}
```

Problem 6:
```{r a6, ref.label='p6_a', eval=F, echo=T}
```

Problem 7:
```{r a7, ref.label='p7_a', eval=F, echo=T}
```

Problem 10:
```{r a10, ref.label='p10_a', eval=F, echo=T}
```

Problem 11:
```{r a11, ref.label='p11_a', eval=F, echo=T}
```

Problem 12:
```{r a12, ref.label='p12_a', eval=F, echo=T}
```
